---
title: "ESS data exploration and univariate analysis"
author: "Bradley McKenzie"
date: "June 2025"
output:
  html_document: 
    theme: cerulean
    highlight: tango
    toc: yes
    toc_float: true
    toc_depth: 3
    df_print: paged
editor_options: 
  markdown: 
    wrap: 72

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo=T, knitr.purl.inline = FALSE)

```

## Purpose:

This script will run the descriptive statistics for all of the trust variables within the ESS dataset. There are 7 trust variables that are being explored as potential targets in this analysis. This dataset can take some time to run as we apply weightings throughout and are looping over each wave and country of data. In the report, only the data on trust in the UK parliament and EU Parliament are used, but this script creates summary summary statistics, for all rounds and countries.

The plots area only made for select countries and variables of interest: 
 - UK & Germany (study country + comparison for diff-in-diff)
 - trust in UK Parliament, trust in EU parliament

When doing the initial international comparisons, we primarily use unweighted data to make comparison faster, however weights are introduced later for the the final calculations. Weights are applied in all calculations for the UK only. 

The order of this script is: 
1. Call the `Dataframe creation_ESS` script to load ESS data with select key variables
2. Explore the different weights & stratum in ESS - justifying the use of `anweight`.
3. Create functions to create weighted trust values for all variables in all rounds in all countries. 
4. Plot selected trends. 
5. Create variables of extreme high/low trust in different values to monitor for trends (note, this is not used in the final report)
6. Create functions to plot trust variables between the UK and other countries
7. Conduct international comparisons to check for a country-level comparison to the UK

First, run the data cleaning script over the ESS download. Run this first and link to your ESS file download. 

## Read data, load libraries

Load the dataframe creation code, we call the dataframe creation file "Dataframe creation_ESS.Rmd" which reads our datafile in and performs variable selection. This code should be adjusted first to set your data location. 

This file also calculates 'anweight' values for each round of the data.

```{r load-data-silently, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
knitr::purl("Dataframe creation_ESS.Rmd", output = "temp_data_prep.R")
source("temp_data_prep.R")
file.remove("temp_data_prep.R")
```

Load additional libraries:

```{r}
library(survey)
library(srvyr)
library(ggplot2)
library(purrr)
library(forcats)
library(ggrepel)
```

## Understanding survey weights in ESS and logic for simple 'anweight' use.

This section performs checks on the weighting objects within ESS and provides justification for why we use the anweight as our basic adjustment across each round of data rather than choosing the combined simple and complex weighting approaches. 

From the ESS Methodology Report:

"For rounds 1-8, no PSU information is available. Users should be aware that not accounting for clustering may lead to underestimated standard errors."

      For ESS Rounds 1-8 (no psu/stratum):
      Use a pseudo-design acknowledging limitations:
      
      ess1_8_design <- svydesign(
        ids = ~1, # No clustering info available
        weights = ~anweight,
        data = ess_data_round1_8
      )

Limited because: 
  Clearly document this assumes SRS (simple random sampling).
  Interpret SEs cautiously: they're underestimated.
  Avoid bold claims about statistical significance.

But we see we are missing the key vars, psu and stratum.

To overcome this, we would need to integrate many country level datsets and calculate weights which is not feasible in the timeframes. 

```{r}
# Show missing weights data until R8
ess_data |> group_by(essround) |> summarise(total_psu = sum(psu, na.rm = TRUE),
                                            total_stratum = sum(stratum, na.rm=TRUE))

```

"From round 9 onwards, psu and stratum are included. Use these for proper variance estimation."

For ESS Rounds 9-11 (with psu and stratum):
    ess9_11_design <- svydesign(
      ids = ~psu, 
      strata = ~stratum, 
      weights = ~anweight,
      data = ess_data_round9_11
    )

So we create the 2 samples, then we can compare our SRS approach and the weighted approach from wave 9 onwards to check estimates. 

```{r}
    # create data subset first for r9, 10 and 11 modelling 
ess_data_r9_r11 <- ess_data |> filter(essround %in% c(9,10,11), cntry == "GB")

```

## Create a survey design object(s)

```{r, include }
## complete, SRS model with primary sampling unit = 1 
simple_design <- svydesign(ids = ~1 , 
                           data = ess_data, 
                           weights = ~anweight)

# more detailed weighted model for w9-11
complex_design_r9_r11 <- svydesign(ids = ~psu, 
                        strata = ~stratum, 
                        weights = ~anweight, 
                        data = ess_data_r9_r11,
                        nest = TRUE)

```


Now, calculate the mean values in two central trust variables to compare differences in estimation method over time: 

Note that the complex weighting method takes significantly longer to run as well so only test if you want to confirm the findings. 

```{r, inlcude=FALSE}
simple_summary_data <- svyby(
  formula = ~trust_people,
  by = ~essround + cntry,
  design = simple_design,
  FUN = svymean,
  na.rm = TRUE,
  vartype = "ci", # To get confidence intervals
  ci.level = 0.95 # 95% confidence interval
)

complex_summary_data <- svyby(
  formula = ~trust_people,
  by = ~cntry + essround,
  design = complex_design_r9_r11,
  FUN = svymean,
  na.rm = TRUE,
  vartype = "ci", # To get confidence intervals
  ci.level = 0.95 # 95% confidence interval
)

simple_summary_data
complex_summary_data
```

Then we compare the data for just R9,10 & 11 for the UK to get an idea of what errors we may experience with variance

```{r}
simple_summary_data |> filter(essround %in% c(9,10,11)) |> 
  left_join(complex_summary_data, by = c("cntry", "essround"), suffix = c(".simple", ".complex")) |> 
  mutate(diff_avg = trust_people.simple - trust_people.complex,
         diff_lower = round(ci_l.simple - ci_l.complex,4),
         diff_upper = round(ci_u.simple - ci_u.complex,4)) |> 
  select(essround, cntry, starts_with("diff"))
  
```

We see through these that our estimates are identical, but we do lose some precision in the variance of our estiamtes as flagged in the weighting documentation. 

We will note this in our limitations but note it does not impact our point estimates.  

## Weighted univariate analysis for all trust variables:

Here, we create the average score for all trust variables for all years and countries with confidence intervals, then join the data together.

A function is created to calculate the weighted average for each varaible which is looped through.  

```{r}

# Get all variable names that start with "trust"
trust_vars <- names(simple_design$variables)[grepl("^trust", names(simple_design$variables))]

# create function to run the summary for each
trust_summary_simple <- function(var) {
  formula_str <- paste0("~", var)
  
  svyby(
    formula = as.formula(formula_str),
    by = ~cntry + essround,
    design = simple_design,
    FUN = svymean,
    vartype = "ci",
    ci.level = 0.95,
    na.rm = TRUE
  ) %>%
    mutate(variable = var)  # Add variable name for identification/standardisation
}

# run mapdfr to get a df and loop the summary functions over each trust variable
trust_summaries_df <- map_dfr(trust_vars, trust_summary_simple)
print(trust_summaries_df)

# now join all the trust_columns into one with pivot longer and tidy the df. 
trust_summaries_df <- trust_summaries_df |>
  select(-variable) |> # remove duplicated col
  pivot_longer(
    cols = starts_with("trust"),  
    names_to = "trust_variable",       
    values_to = "wt_avg_mean"    
  ) |> 
  drop_na() |>  # introduces many NA columns, so we remove for each 
  select(essround, cntry, trust_variable, wt_avg_mean, ci_l, ci_u) #reorder data

print(trust_summaries_df)
```

We now have a clean dataset with the average trust levels for each wave, adjusted for anweight, in each survey round, for each country. 

We can see that trust in scientists was only available in 

We check the distribution and availability of each variable:

We see that trust in science wasonly asked in round 10. So we remove this. 
Trust in political parties was not asked in Round 1. But we can impute this data as NA. 
```{r}
# check the missing values -> where the data == 0
trust_summaries_df |> 
  filter(wt_avg_mean == 0) |> 
  group_by(trust_variable) |> 
  summarise(total_missing = n(),
            total_rounds = length(unique(essround)))

# check the total valid responses by country per round
trust_summaries_df |> 
  group_by(essround, trust_variable) |> 
  summarise(total_valid = sum(wt_avg_mean>0)) |> 
  pivot_wider(names_from = trust_variable,
              values_from = total_valid)


# Remove trust in science and convert 0 to NA
trust_summaries_df <- trust_summaries_df |> 
  filter(trust_variable != "trust_scien") |> # remove science
  mutate(across(where(is.numeric), ~na_if(., 0))) # convert to NA
```


### Now plot all UK trends first

First create a data subset, this will be used later to compare vs other countries too 

```{r}
gb_trust_summaries <- trust_summaries_df |> filter(cntry == "GB")

```


And plot - all trends for UK: 

```{r}
# Create the line chart with error bars
ggplot(gb_trust_summaries, 
       aes(x = essround, y = wt_avg_mean, 
           group = trust_variable, colour = trust_variable)) +  # group=1 ensures lines connect across points
  geom_line(linewidth = 1) +  # Line for the trend
  geom_point(size = 1.5) +  # Points for each mean
  geom_errorbar(aes(ymin = ci_l, ymax = ci_u),  # Error bars using CI bounds
                width = 0.5,  # Width of the error bar caps
                color = "gray") +  
  labs(title = "Mean Trust in UK by variable, ESS round 1-11",
       subtitle = "With 95% Confidence Intervals",
       x = "ESS Round",
       y = "Mean Trust (/10 with CI included)") +
  theme_minimal()+
  
  # add extra labels to see each item
  geom_label_repel(
    data = . %>% group_by(trust_variable) %>% filter(essround == max(essround)),
    aes(label = trust_variable, fill = trust_variable), 
    color = "white",                # <--- CHANGE 1: Set text color inside labels to white
    fontface = "bold",
    alpha = 0.8,
    label.size = 0.5,
    show.legend = FALSE,
    box.padding = unit(0.5, "lines"),
    point.padding = unit(0.5, "lines"),
    segment.colour = NA,
    direction = "y",
    nudge_x = 2,
    min.segment.length = 0)+
    # set limits 
  scale_x_continuous(breaks = seq(1,11,1),
                     limits = c(min(gb_trust_summaries$essround), max(gb_trust_summaries$essround)+3)) +
  geom_vline(xintercept = 7.5, color = "orange", size = 1.5, alpha = 0.5)+
  theme(legend.position = "none")



```

Facet for UK only to have each variable separate 
```{r}
# And print faceted by country too
ggplot(gb_trust_summaries, 
       aes(x = factor(essround), y = wt_avg_mean, 
           group = trust_variable, colour = trust_variable)) +  # group=1 ensures lines connect across points
  geom_line(size = 1) +  # Line for the trend
  geom_point(size = 1.5) +  # Points for each mean
  geom_errorbar(aes(ymin = ci_l, ymax = ci_u),  # Error bars using CI bounds
                width = 0.5,  # Width of the error bar caps
                color = "gray") +  
  facet_wrap(~trust_variable)+
  labs(title = "Mean Trust in UK by variable, ESS round 1-11",
       subtitle = "With 95% Confidence Intervals",
       x = "ESS Round",
       y = "Mean Trust (/10 with CI included)") +
  theme_minimal() + 
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for readability
    legend.position = "none"
  )+
  geom_vline(xintercept = 7.5) # sample of brexit timing

```

### Remove trust in science (missing)

As seen above, these are missing: 

```{r}
ess_data <- ess_data |> 
  select(-trust_scien)
```

We keep R1 of political parties as we can't remove just the one variable. 

## Investigating polarised values of trust

Create variables for very high trust and very low trust, we will group as: 
0,1,2,3: "Low trust"
4,5,6: "Moderate trust 
7,8,9,10: "High trust"

With a second variable we recode as: 
0,1: "Extreme low trust"
9,10: "Extreme high trust"

```{r}
# proof of concept for function to classify high/low
ess_data |> 
  mutate(trust_hilow_people = factor(case_when(trust_people<4 ~ "Low",
                                        between(trust_people, 4,6) ~ "Moderate",
                                        between(trust_people, 7,10) ~ "High",
                                        TRUE ~ NA),
                                     levels = c("Low", "Moderate", "High", NA))) |> 
  group_by(trust_hilow_people) |> count()

# check why NA values included  
ess_data |> 
  mutate(trust_hilow_people = case_when(trust_people<4 ~ "Low",
                                        between(trust_people, 4,6) ~ "Moderate",
                                        between(trust_people, 7,10) ~ "High",
                                        TRUE ~ NA)) |> 
  select(essround, cntry, trust_people, trust_hilow_people) |> 
  filter(is.na(trust_hilow_people))

```

We see that the NA's are true NA's so we can continue with our calculations: 

```{r}
#function to create high/low variables for each: 
low_high_trust <- function(trust_var) {
  factor(
    case_when(
      trust_var < 4 ~ "Low",
      between(trust_var, 4, 6) ~ "Moderate",
      between(trust_var, 7, 10) ~ "High", # use range to avoid any poorly coded 77/88/99 values being included
      TRUE ~ NA #account for true NAs from no response
  ),
  levels = c("High","Moderate","Low",  NA))
}

# and function for extreme trust (top/bottom 2 values)
extreme_trust <- function(trust_var) {
  factor(
    case_when(
      trust_var < 2 ~ "Extreme low",
      between(trust_var, 2, 8) ~ "Moderate",
      between(trust_var, 9, 10) ~ "Extreme high",
      TRUE ~ NA
  ),
  levels = c("Extreme high", "Moderate",  "Extreme low")
  )
}

# now apply over our trust variables to get the trend in trust
ess_data <- ess_data |> 
  mutate(across(matches("^trust") & !matches("hilow|extreme"), low_high_trust, .names = "{.col}_hilow"),
         across(matches("^trust") & !matches("hilow|extreme"), extreme_trust, .names = "{.col}_extreme")) 

```

Now calculate the unweighted rates within each country and wave 

```{r}
# subset the data
gb_polarise_trust <- ess_data |> 
  filter(cntry == "GB") |> 
  select(anweight, essround, cntry, matches("_extreme|_hilow"))

print(gb_polarise_trust)

```

Now calculate the weighted values for the high, moderate and low. 
We reuse the function from above as we are still calculating the means of each group 

```{r}
# Get all variable names that start with "trust"
trust_vars_hilow <- names(gb_polarise_trust)[grepl("*hilow", names(gb_polarise_trust))]

# update the survey design to recognise the variables: 
simple_design_polarise <- svydesign(ids = ~1 , 
                           data = gb_polarise_trust, 
                           weights = ~anweight)

# # create function to run the summary for each - same as above with just different design called
trust_summary_hilow_simple <- function(var) {
  formula_str <- paste0("~", var)

  svyby(
    formula = as.formula(formula_str),
    by = ~cntry + essround,
    design = simple_design_polarise,
    FUN = svymean,
    vartype = "ci",
    ci.level = 0.95,
    na.rm = TRUE
  ) %>%
    mutate(variable = var)  # Add variable name for identification/standardisation
}

# run mapdfr to get a df and loop the summary functions over each trust variable
trust_hilow_summaries_df <- map_dfr(trust_vars_hilow, trust_summary_hilow_simple)
print(trust_hilow_summaries_df)
```

Now clean the data for plotting. 

```{r}
# now join all the trust_columns into one with pivot longer and tidy the df. 
trust_hilow_summaries_df2 <- trust_hilow_summaries_df |>
  select(-variable) |> # remove duplicated col
  pivot_longer(
    cols = starts_with("trust"),  
    names_to = "trust_variable",       
    values_to = "prop_pop"    
  ) |> 
  pivot_longer(
    cols = starts_with("ci_l"),
    names_to = "ci_lower_group",
    values_to = "ci_lower"
  ) |>
  pivot_longer(
    cols = starts_with("ci_u"),
    names_to = "ci_upper_group",
    values_to = "ci_upper"
  ) |>
 drop_na() |>  # introduces many NA columns, so we remove for each 
    # here we filter only for the rows that are matching. It has matched for all possible Ci upper and lower values
  filter(trust_variable == substr(ci_lower_group, 6, nchar(ci_lower_group)) & trust_variable == substr(ci_upper_group, 6, nchar(ci_upper_group))) |> 
  select(essround, cntry, trust_variable, prop_pop, ci_lower, ci_upper) #reorder data

# and split the name and value out. 
trust_hilow_summaries_df2 <- trust_hilow_summaries_df2 |> 
    separate_wider_delim(
      col = trust_variable,
      delim = "_hilow", # The characters to split on
      names = c("trust_variable", "level"), # Names for the new columns
      too_many = "merge" # If there were more parts after "_hilow", merge them into 'level'
    ) 
# and convert variables to leveled factors:
trust_hilow_summaries_df2 <- trust_hilow_summaries_df2 |> 
  mutate(trust_variable = factor(trust_variable),
         level = factor(level, levels = c("Low", "Moderate", "High")))



print(trust_hilow_summaries_df2)
```

We now have weighted data for the proportions of high/moderate/low groups for each country in each round across the dataset. 

Plot high and low trust views for UK in bars:

```{r}
# Now we plot the trends in each. 

trust_hilow_summaries_df2 |>
  ggplot(aes(x = factor(essround), # x-axis for ESS Round
             y = prop_pop,          # Map your pre-calculated proportion to y
             fill = level)) +       # Fill based on the 'level' factor
  geom_bar(position = "stack", stat = "identity", color = "grey20", linewidth = 0.1) + # Stack pre-calculated proportions
                                                                         # Added 'color' for bar outlines
  scale_y_continuous(labels = scales::percent_format()) + # Format y-axis as percentage
  scale_fill_manual(values = c("Low" = "red", "High" = "blue", "Moderate" = "NA")) + # custom colours to hide moderate
  facet_wrap(~trust_variable, ncol = 4) + # Create separate plots for each trust_variable
  labs(title = "Proportion of UK with strong views on trust (0-3 = Low; 7-10 = High), weighted R1-R11",
       x = "ESS Round", y = "% of the UK", fill = "Trust Level") +
  theme_minimal() +
  theme(
    panel.grid.major.x = element_blank(), # Remove vertical grid lines 
    panel.grid.minor.x = element_blank()
  )

```

### Changes in extreme high/low trust (top/bottom 2 scores.)
And with the extreme trust levels (0-1 or 9-10):

```{r}
# pivot longer for ggploting
gb_polarise_trust |> 
  select(essround, ends_with("extreme"), -matches("scien")) |> 
  pivot_longer(cols = ends_with("extreme"), 
               names_to = "trust_type", 
               values_to = "trust_category") |> 
  drop_na() |> # remove science and R1 parties
    # then plot stacked bars
  ggplot(aes(x = factor(essround), 
             fill = trust_category)) + 
  geom_bar(position = "fill") +  # stack to 100%
  scale_y_continuous(labels = scales::percent_format()) +  
  scale_fill_manual(values = c("Extreme low" = "red", "Extreme high" = "blue", "Moderate" = "NA")) +  # custom colours to hide moderate
  facet_wrap(~trust_type, ncol=4) +  # Separate plots 
  labs(title = "Extreme trends in trust vars (0-1 = Low; 9-10 = High)",
       x= "ESS Round", y = "respondents %", fill = "Trust Level") +
  theme_minimal()
```

#### Comparing local and european parliament trust:

```{r}
# Create the line chart with error bars
gb_trust_summaries |> 
  filter(trust_variable %in% c("trust_europ", "trust_parliament")) |> 
ggplot(aes(x = essround, y = wt_avg_mean, 
           group = trust_variable, colour = trust_variable)) +  # group=1 ensures lines connect across points
  geom_line(size = 1) +  # Line for the trend
  geom_point(size = 1.5) +  # Points for each mean
  geom_errorbar(aes(ymin = ci_l, ymax = ci_u),  # Error bars using CI bounds
                width = 0.5,  # Width of the error bar caps
                color = "gray") +  
  labs(title = "Mean Trust in Parliament - UK vs European, ESS round 1-11",
       subtitle = "With 95% Confidence Intervals",
       x = "ESS Round",
       y = "Mean Trust (/10 with CI included)") +
  theme_minimal()+
  
  # add extra labels to see each item
  geom_label_repel(
    data = . %>% group_by(trust_variable) %>% filter(essround == max(essround)),
    aes(label = trust_variable, fill = trust_variable), 
    color = "white",                # <--- CHANGE 1: Set text color inside labels to white
    fontface = "bold",
    alpha = 0.8,
    label.size = 0.5,
    show.legend = FALSE,
    box.padding = unit(0.5, "lines"),
    point.padding = unit(0.5, "lines"),
    segment.colour = NA,
    direction = "y",
    nudge_x = 2,
    min.segment.length = 0)+
    # set limits 
  scale_x_continuous(breaks = seq(1,11,1),
                     limits = c(min(gb_trust_summaries$essround), max(gb_trust_summaries$essround)+3)) +
  geom_vline(xintercept = 7.5, color = "orange", size = 1.5, alpha = 0.5)+
  theme(legend.position = "none")


```


### International comparisons: 

```{r}
trust_summaries_df |> 
  ggplot(aes(x=essround,
             y=wt_avg_mean,
             group=cntry,
             colour = cntry))+
  geom_line()+
  # set limits 
  scale_x_continuous(breaks = seq(1,11,1),
                     limits = c(min(trust_summaries_df$essround), max(trust_summaries_df$essround)+1)) +
  theme_minimal()+
  geom_vline(xintercept = 7.5, color = "orange", size = 1.5, alpha = 0.5)+
  labs(title = "Average trust in [...] - ESS rounds 4 to 11, UK only",
     x = "ESS round", y = "Mean trust (0‒10)")+
  theme(legend.position = "bottom")+
  facet_wrap(~trust_variable)

```


#### Comparing individual countries

Matching in ESS_trust_w678.Rmd. The section is just for visual inspections. 

First, testing plots of just one country:

Create plot to test parallel trends between UK and other country. We write a function to complete the plotting against multiple countries more easily. 

```{r}
## print sample for comparing by country. 
trust_summaries_df |> 
  filter(cntry %in% c("GB", "DE")) |> 
  ggplot(aes(x = factor(essround), # Treat essround as a factor for discrete x-axis labels
             y = wt_avg_mean,
             group = cntry, # Group by country to draw lines for each country
             colour = cntry)) +
  geom_line(linewidth = 1) + # Added linewidth for better visibility
  geom_point(size = 2) + # Add points for each data point
  geom_vline(xintercept = 7.5, color = "orange", size = 1.5, alpha = 0.5)+
  facet_wrap(~trust_variable) +
  theme_minimal()+
  theme(legend.position = "bottom")+
  labs(title = "Average trust in [...] - ESS Round 1-11, GB vs DE",
   x = "ESS round", 
   y = "Mean trust (0‒10)",
   colour = "Country:")
```


Now make the function to plot GB vs other countries 
```{r}
# now create as function:
plot_gb_other <- function(summary_stats= summary_stats, other_country){
  summary_stats |> 
    select(essround, cntry, 
         people = ppltrst_mean, 
         euro_parl = trstep_mean, 
         legal_sys = trstlgl_mean,
         police = trstplc_mean,
         politicians = trstplt_mean,
         parliament = trstprl_mean,
         polit_parties = trstprt_mean,
         united_nats = trstun_mean
         ) |> 
  pivot_longer(cols = people:united_nats) |> 
  filter(cntry %in% c("GB", other_country),
         essround %in% c(6,7)) |> 
  ggplot(aes(x = factor(essround), # Treat essround as a factor for discrete x-axis labels
             y = value,
             group = cntry, # Group by country to draw lines for each country
             colour = cntry)) +
  geom_line(linewidth = 1) + # Added linewidth for better visibility
  geom_point(size = 2) + # Add points for each data point
  geom_vline(xintercept = 7.5, color = "orange", size = 1.5, alpha = 0.5)+
  facet_wrap(~name, scales = "free_y") +
  theme_minimal()+
  theme(legend.position = "bottom")+
  labs(title = glue("Average trust in [...] - ESS Round 6-7 (baseline), GB vs {other_country}"),
   x = "ESS round", 
   y = "Mean trust (0‒10)",
   colour = "Country:")
}

trust_summaries_df |> 
  filter(cntry %in% c("GB", "DE")) |> 
  ggplot(aes(x = factor(essround), # Treat essround as a factor for discrete x-axis labels
             y = wt_avg_mean,
             group = cntry, # Group by country to draw lines for each country
             colour = cntry)) +
  geom_line(linewidth = 1) + # Added linewidth for better visibility
  geom_point(size = 2) + # Add points for each data point
  geom_vline(xintercept = 7.5, color = "orange", size = 1.5, alpha = 0.5)+
  facet_wrap(~trust_variable) +
  theme_minimal()+
  theme(legend.position = "bottom")+
  labs(title = "Average trust in [...] - ESS Round 1-11, GB vs DE",
   x = "ESS round", 
   y = "Mean trust (0‒10)",
   colour = "Country:")


# now create as function:
plot_gb_other_parliament <- function(trust_summaries_df=trust_summaries_df, other_country){
  trust_summaries_df |> 
    filter(cntry %in% c("GB", other_country),
           trust_variable %in% c("trust_europ", "trust_parliament")) |> 
  ggplot(aes(x = factor(essround), # Treat essround as a factor for discrete x-axis labels
             y = wt_avg_mean,
             group = cntry, # Group by country to draw lines for each country
             colour = cntry)) +
  geom_line(linewidth = 1) + # Added linewidth for better visibility
  geom_point(size = 2) + # Add points for each data point
  geom_vline(xintercept = 7.5, color = "orange", size = 1.5, alpha = 0.5)+
  facet_wrap(~trust_variable, scales = "free_y") +
  theme_minimal()+
  theme(legend.position = "bottom")+
  labs(title = glue("Average trust in EU and local Parliament - ESS Round 1-11, GB vs {other_country}"),
   x = "ESS round", 
   y = "Mean trust (0‒10)",
   colour = "Country:")
}

plot_gb_other_parliament(trust_summaries_df, "DE")
```

```{r}

```


Test plot:

```{r}
trust_summaries_df |> 
  filter(trust_variable %in% c("trust_europ", "trust_parliament"),
         cntry %in% c("GB", "DE")) |> 
ggplot(aes(x = essround, y = wt_avg_mean, 
           group = cntry, colour = cntry)) +  
  geom_line(size = 1) +  # Line for the trend
  geom_point(size = 1.5) +  # Points for each mean
  geom_errorbar(aes(ymin = ci_l, ymax = ci_u),  # Error bars using CI bounds
                width = 0.5,  # Width of the error bar caps
                color = "gray") +  
  labs(title = "Mean Trust in Parliament - UK vs European, ESS round 1-11",
       subtitle = "With 95% Confidence Intervals",
       x = "ESS Round",
       y = "Mean Trust (/10 with CI included)") +
  theme_minimal()+
  
  # add extra labels to see each item
  geom_label_repel(
    data = . %>% group_by(trust_variable) %>% filter(essround == max(essround)),
    aes(label = cntry, fill = cntry), 
    color = "white",    
    fontface = "bold",
    alpha = 0.8,
    label.size = 0.5,
    show.legend = FALSE,
    box.padding = unit(0.5, "lines"),
    point.padding = unit(0.5, "lines"),
    segment.colour = NA,
    direction = "y",
    nudge_x = 2,
    min.segment.length = 0)+
    # set limits 
  scale_x_continuous(breaks = seq(1,11,1),
                     limits = c(min(trust_summaries_df$essround), max(trust_summaries_df$essround)+3)) +
  geom_vline(xintercept = 7.5, color = "orange", size = 1.5, alpha = 0.5)+
  theme(legend.position = "none")+
  facet_wrap(~trust_variable)



```

### Parallel trends plot UK vs Germany

```{r}
### plotting just the pre-Referendum periods 
trust_summaries_df |> 
  # limit data 
  filter(trust_variable == "trust_europ",
         cntry %in% c("GB", "DE"),
         essround <9) |> 
  #clean names for plotting
  mutate(cntry = ifelse(cntry == "GB", "UK", "Germany"),
         trust_variable = ifelse(trust_variable == "trust_europ", "Trust in European Parliament", trust_variable)) |> 
  # plot 
ggplot(aes(x = essround, y = wt_avg_mean, 
           group = cntry, colour = cntry)) +  
  geom_line(size = 1) +  # Line for the trend
  geom_point(size = 1.5) +  # Points for each mean
  geom_errorbar(aes(ymin = ci_l, ymax = ci_u),  # Error bars using CI bounds
                width = 0.5,  # Width of the error bar caps
                color = "gray") +  
  labs(title = "Mean Trust in Parliament - UK vs Germany, ESS round 1-8",
       subtitle = "With 95% Confidence Intervals",
       x = "ESS Round",
       y = "Mean Trust (weighted)") +
  theme_minimal()+
  
  # add extra labels to see each item
  geom_label_repel(
    data = . %>% group_by(trust_variable) %>% filter(essround == max(essround)),
    aes(label = cntry, fill = cntry), 
    color = "white",    
    fontface = "bold",
    alpha = 0.8,
    label.size = 0.5,
    show.legend = FALSE,
    box.padding = unit(0.5, "lines"),
    point.padding = unit(0.5, "lines"),
    segment.colour = NA,
    direction = "y",
    nudge_x = 0.75,
    min.segment.length = 0)+
    # set limits 
  scale_x_continuous(breaks = seq(1,8,1),
                     limits = c(min(trust_summaries_df$essround),9)) +
  # add referendum line
  geom_vline(xintercept = 7.7, color = "orange", size = 1.5, alpha = 0.5)+
  annotate(geom="text", x = 7.4, y =4.5, label = "Referendum",
             angle = 90, color = "orange")+
  theme(legend.position = "none",
        panel.grid.major.x = element_blank(), # Remove vertical grid lines 
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()
  )
```

### UK vs Germany with standardised values to R1
```{r}
### plotting just the pre-Referendum periods 
trust_summaries_df |> 
  # limit data 
  filter(trust_variable == "trust_europ",
         cntry %in% c("GB", "DE"),
         essround <9) |> 
  #clean names for plotting
  mutate(cntry = ifelse(cntry == "GB", "UK", "Germany"),
         trust_variable = ifelse(trust_variable == "trust_europ", "Trust in European Parliament", trust_variable)) |> 
  group_by(cntry) |> 
  
# now standardise the values to R1 
    mutate(R1_baseline = wt_avg_mean[essround == 1],
           wt_avg_mean = wt_avg_mean - R1_baseline,
           ci_l = ci_l - R1_baseline,
           ci_u = ci_u - R1_baseline) |> 
  ungroup() |> 
  
# plot 
ggplot(aes(x = essround, y = wt_avg_mean, 
           group = cntry, colour = cntry)) +  
  geom_line(size = 1) +  # Line for the trend
  geom_point(size = 1.5) +  # Points for each mean
  geom_errorbar(aes(ymin = ci_l, ymax = ci_u),  # Error bars using CI bounds
                width = 0.5,  # Width of the error bar caps
                color = "gray") +  
  labs(title = "Mean Trust in Parliament - UK vs Germany, ESS round 1-8",
       subtitle = "With 95% Confidence Intervals",
       x = "ESS Round",
       y = "Mean Trust (weighted)") +
  theme_minimal()+
  
  # add extra labels to see each item
  geom_label_repel(
    data = . %>% group_by(trust_variable) %>% filter(essround == max(essround)),
    aes(label = cntry, fill = cntry), 
    color = "white",    
    fontface = "bold",
    alpha = 0.8,
    label.size = 0.5,
    show.legend = FALSE,
    box.padding = unit(0.5, "lines"),
    point.padding = unit(0.5, "lines"),
    segment.colour = NA,
    direction = "y",
    nudge_x = 0.75,
    min.segment.length = 0)+
    # set limits 
  scale_x_continuous(breaks = seq(1,8,1),
                     limits = c(min(trust_summaries_df$essround),9)) +
  # add referendum line
  geom_vline(xintercept = 7.7, color = "orange", size = 1.5, alpha = 0.5)+
  annotate(geom="text", x = 7.4, y =0, label = "Referendum",
             angle = 90, color = "orange")+
  theme(legend.position = "none",
        panel.grid.major.x = element_blank(), # Remove vertical grid lines 
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()
  )
```


### Other country comparisons 


Now turn to function to compare other countries 

```{r}
plot_gb_other_parliament <- function(trust_summaries_df=trust_summaries_df, other_country){
    
  trust_summaries_df |> 
      filter(trust_variable %in% c("trust_europ", "trust_parliament"),
             cntry %in% c("GB", other_country)) |> 
    ggplot(aes(x = essround, y = wt_avg_mean, 
               group = cntry, colour = cntry)) +  
      geom_line(size = 1) +  # Line for the trend
      geom_point(size = 1.5) +  # Points for each mean
      geom_errorbar(aes(ymin = ci_l, ymax = ci_u),  # Error bars using CI bounds
                    width = 0.5,  # Width of the error bar caps
                    color = "gray") +  
      labs(title =  glue("Average trust in EU and local Parliament - ESS Round 1-11, GB vs {other_country}"),
           subtitle = "With 95% Confidence Intervals",
           x = "ESS Round",
           y = "Mean Trust") +
      theme_minimal()+
  
  # add extra labels to see each item
  geom_label_repel(
    data = . %>% group_by(trust_variable) %>% filter(essround == max(essround)),
    aes(label = cntry, fill = cntry), 
    color = "white",    
    fontface = "bold",
    alpha = 0.8,
    label.size = 0.5,
    show.legend = FALSE,
    box.padding = unit(0.5, "lines"),
    point.padding = unit(0.5, "lines"),
    segment.colour = NA,
    direction = "y",
    nudge_x = 2,
    min.segment.length = 0)+
    # set limits 
  scale_x_continuous(breaks = seq(1,11,1),
                     limits = c(min(trust_summaries_df$essround), max(trust_summaries_df$essround)+3)) +
  geom_vline(xintercept = 7.5, color = "orange", size = 1.5, alpha = 0.5)+
  theme(legend.position = "none")+
  facet_wrap(~trust_variable)
}

```

## Compare parliament outcomes for GB and individual countries:
Here we can compare countries to look at comparability for parallel trends estimates. 

Countries in the dataset: 

```{r}
ess_data |> group_by(cntry, country_name) |> count()

```


Germany (DE):

```{r}
plot_gb_other_parliament(trust_summaries_df, "DE")
```

Ireland (IE):

```{r}
plot_gb_other_parliament(trust_summaries_df, "IE")

```

Belgium (BE):
- doens't look feasible for parallel trends. 
```{r}
plot_gb_other_parliament(trust_summaries_df, "BE")
```

Switzerland:
```{r}
plot_gb_other_parliament(trust_summaries_df, "CH")
```

France:
- Also doesn't look promising. Trust in local parliament is massively differnt in the lead up to brexit. 
```{r}
plot_gb_other_parliament(trust_summaries_df, "FR")
```


Netherlands: 

```{r}
plot_gb_other_parliament(trust_summaries_df, "NL")

```

Scandi countries: 
The trends appear too positive pre-Brexit ot be comaprable. 
```{r}
plot_gb_other_parliament(trust_summaries_df, c("FI", "NO", "SE"))

```

Remaining EU countries: 
Not many promiising options here! 
```{r}
plot_gb_other_parliament(trust_summaries_df, "ES")
plot_gb_other_parliament(trust_summaries_df, "LT")
plot_gb_other_parliament(trust_summaries_df, "CZ")
plot_gb_other_parliament(trust_summaries_df, "HU")
plot_gb_other_parliament(trust_summaries_df, "PL")
plot_gb_other_parliament(trust_summaries_df, "PT")
plot_gb_other_parliament(trust_summaries_df, "SI")
plot_gb_other_parliament(trust_summaries_df, "EE")
```

Germany appears the best to satisfy parallel trends for a comparison of data. 



### Trends for overall EU Parliament trust vs UK parliament - correlation over time by trust group. 

```{r}
ess_data |> 
  filter(cntry == "GB") |> 
  group_by(essround, trust_europ_hilow) |>  
  count() |> pivot_wider(names_from = trust_europ_hilow, values_from = n)

# see which groups are the most volatile. We group by high 
ess_data |> 
  filter(cntry == "GB") |> 
  group_by(essround, trust_europ_hilow) |>  
  select(trust_europ, trust_parliament) |> 
  drop_na() |> 
  summarise(coef = cor(trust_europ, trust_parliament)) |> 
  ggplot(aes(x=essround, y=coef, colour = trust_europ_hilow))+
  geom_point()+geom_line()
```


## Create summaries of trust in europe variables for all countries in each round. 

Here we calculate the mean values of trust in each of the trust variables

### Plot the individual trends 

Create summary data:

```{r}

simple_summary_data <- svyby(
  formula = ~trust_europ,
  by = ~cntry + essround,
  design = simple_design,
  FUN = svymean,
  vartype = "ci", # To get confidence intervals
  ci.level = 0.95, # 95% confidence interval
  na.rm = TRUE # Remove NA observations
)

```

And plot the trend over time with error bars for each year:

### Updated Code for Line Chart with Error Bars

Here's the updated ggplot code for trust in people. 

```{r}

# Create the line chart with error bars, faceted by country
ggplot(simple_summary_data, 
       aes(x = factor(essround), y = trust_europ, group = 1)) +  # group=1 ensures lines connect across points
  geom_line(color = "blue", size = 1) +  # Line for the trend
  geom_point(color = "blue", size = 2) +  # Points for each mean
  geom_errorbar(aes(ymin = ci_l, ymax = ci_u),  # Error bars using CI bounds
                width = 0.5,  # Width of the error bar caps
                color = "gray") +  
  facet_wrap(~cntry) +  # Facet by country; use free_y for independent y-scales
  labs(title = "Mean Trust in People Over Time by Country",
       subtitle = "With 95% Confidence Intervals",
       x = "ESS Round",
       y = "Mean Trust in EU Parliament by country") +
  theme_minimal() +  # Clean theme
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for readability
    strip.text = element_text(size = 12, face = "bold")  # Bold facet labels
  )
```


Trust in politics
