---
title: "Univariate analysis"
author: "bmck"
date: "2025-06-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Purpose:

This script will 

First, run the data cleaning script over the ESS download. Run this first and link to your ESS file download. 

```{r}
### INSERT CODE TO RUN A DATA CLEANING SCRIPT THAT WILL BE KEPT SEPARATE
```

Load additional libraries:

```{r}
library(survey)
library(srvyr)
library(ggplot2)
library(purrr)
library(forcats)
```

## Understanding survey weights in ESS

From the ESS Methodology Report:

"For rounds 1-8, no PSU information is available. Users should be aware that not accounting for clustering may lead to underestimated standard errors."

      For ESS Rounds 1-8 (no psu/stratum):
      Use a pseudo-design acknowledging limitations:
      
      ess1_8_design <- svydesign(
        ids = ~1, # No clustering info available
        weights = ~anweight,
        data = ess_data_round1_8
      )

Limited because: 
  Clearly document this assumes SRS (simple random sampling).
  Interpret SEs cautiously: they're underestimated.
  Avoid bold claims about statistical significance.

But we see we are missing the key vars, psu and stratum.

To overcome this, we would need to integrate many country level datsets and calculate weights which is not feasible in the timeframes. 

```{r}
# Show missing weights data until R8
ess_data |> group_by(essround) |> summarise(total_psu = sum(psu, na.rm = TRUE),
                                            total_stratum = sum(stratum, na.rm=TRUE))

```


"From round 9 onwards, psu and stratum are included. Use these for proper variance estimation."

For ESS Rounds 9-11 (with psu and stratum):
    ess9_11_design <- svydesign(
      ids = ~psu, 
      strata = ~stratum, 
      weights = ~anweight,
      data = ess_data_round9_11
    )

So we create the 2 samples, then we can compare our SRS approach and the weighted approach from wave 9 onwards to check estimates. 
```{r}
    # create data subset first for r9, 10 and 11 modelling 
ess_data_r9_r11 <- ess_data |> filter(essround %in% c(9,10,11))

```



# Create a survey design object(s)

```{r}
## complete, SRS model with primary sampling unit = 1 
simple_design <- svydesign(ids = ~1 , 
                           data = ess_data, 
                           weights = ~anweight)

# more detailed weighted model for w9-11
complex_design_r9_r11 <- svydesign(ids = ~psu, 
                        strata = ~stratum, 
                        weights = ~anweight, 
                        data = ess_data_r9_r11,
                        nest = TRUE)

```


Now, calculate the mean values in two central trust variables to compare differences over time: 

```{r}
simple_summary_data <- svyby(
  formula = ~trust_people,
  by = ~cntry + essround,
  design = simple_design,
  FUN = svymean,
  vartype = "ci", # To get confidence intervals
  ci.level = 0.95 # 95% confidence interval
)

simple_summary_data <- svyby(
  formula = ~trust_people,
  by = ~cntry + essround,
  design = simple_design,
  FUN = svymean,
  vartype = "se", 
  ci.level = 0.95,
  na.rm = TRUE,  # Remove NA observations
  keep.var = TRUE,  # Keep variance estimates
  keep.names = TRUE
)

simple_summary_data
```

```{r}
library(srvyr)
ess_data |> 
  as_survey_design(ids = 1, weights = anweight) |> 
  group_by(cntry, essround) |> 
  summarise(mean_trust = survey_mean(trust_people),
            se_trust = survey_mean(trust_people))

ess_data |> 
  as_survey_design(ids = 1, weights = anweight) |> 
  filter(!is.na(trust_people)) %>%  # Pre-filter missing values
  srvyr::group_by(cntry, essround) %>%
  srvyr::summarize(
    mean_trust = survey_mean(trust_people, na.rm = TRUE),
    se = survey_mean(trust_people, vartype = "se")
  )


```


### Now plot the UK trends

```{r}

```


## Create trust variables for means

Create function to plot trust variables: 


```{r}

summary_data <- svyby(
  formula = ~response_variable,
  by = ~time_period,
  design = design_wt,
  FUN = svymean,
  vartype = "ci", # To get confidence intervals
  ci.level = 0.95 # 95% confidence interval
)

```


### Plot the individual trends 

Create summary data:

```{r}

simple_summary_data <- svyby(
  formula = ~trust_europ,
  by = ~cntry + essround,
  design = simple_design,
  FUN = svymean,
  vartype = "ci", # To get confidence intervals
  ci.level = 0.95, # 95% confidence interval
  na.rm = TRUE # Remove NA observations
)

```

And plot the trend over time with error bars for each year:

### Updated Code for Line Chart with Error Bars

Here's the updated ggplot code for trust in people. 

```{r}

# Create the line chart with error bars, faceted by country
ggplot(simple_summary_data, 
       aes(x = factor(essround), y = trust_europ, group = 1)) +  # group=1 ensures lines connect across points
  geom_line(color = "blue", size = 1) +  # Line for the trend
  geom_point(color = "blue", size = 2) +  # Points for each mean
  geom_errorbar(aes(ymin = ci_l, ymax = ci_u),  # Error bars using CI bounds
                width = 0.5,  # Width of the error bar caps
                color = "gray") +  
  facet_wrap(~cntry, scales = "free_y") +  # Facet by country; use free_y for independent y-scales
  labs(title = "Mean Trust in People Over Time by Country",
       subtitle = "With 95% Confidence Intervals",
       x = "ESS Round",
       y = "Mean Trust in EU Parliament by country") +
  theme_minimal() +  # Clean theme
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for readability
    strip.text = element_text(size = 12, face = "bold")  # Bold facet labels
  )
```


Trust in politics

## Create averages for all years and countries with confidence intervals 

Here we create a function to calculate the weighted average for all of the 

```{r}

# Get all variable names that start with "trust"
trust_vars <- names(simple_design$variables)[grepl("^trust", names(simple_design$variables))]

# create function to run the summary for each
trust_summary_simple <- function(var) {
  formula_str <- paste0("~", var)
  
  svyby(
    formula = as.formula(formula_str),
    by = ~cntry + essround,
    design = simple_design,
    FUN = svymean,
    vartype = "ci",
    ci.level = 0.95,
    na.rm = TRUE
  ) %>%
    mutate(variable = var)  # Add variable name for identification/standardisation
}

# run mapdfr to get a df and loop the summary functions over each trust variable
trust_summaries_df <- map_dfr(trust_vars, trust_summary_simple)
print(trust_summaries_df)

# now join all the trust_columns into one with pivot longer and tidy the df. 
trust_summaries_df <- trust_summaries_df |>
  select(-variable) |> # remove duplicated col
  pivot_longer(
    cols = starts_with("trust"),  
    names_to = "trust_variable",       
    values_to = "wt_avg_mean"    
  ) |> 
  drop_na() |>  # introduces many NA columns, so we remove for each 
  select(essround, cntry, trust_variable, wt_avg_mean, ci_l, ci_u) #reorder data

print(trust_summaries_df)
```

We now have a clean dataset with the average trust levels for each wave, adjusted for anweight, in each survey round, for each country. 

We can see that trust in scientists was only available in 

We check the distribution and availability of each variable:

We see that trust in science wasonly asked in round 10. So we remove this. 
Trust in political parties was not asked in Round 1. But we can impute this data as NA. 
```{r}
# check the missing values -> where the data == 0
trust_summaries_df |> 
  filter(wt_avg_mean == 0) |> 
  group_by(trust_variable) |> 
  summarise(total_missing = n(),
            total_rounds = length(unique(essround)))

# check the total valid responses by country per round
trust_summaries_df |> 
  group_by(essround, trust_variable) |> 
  summarise(total_valid = sum(wt_avg_mean>0)) |> 
  pivot_wider(names_from = trust_variable,
              values_from = total_valid)


# Remove trust in science and convert 0 to NA
trust_summaries_df <- trust_summaries_df |> 
  filter(trust_variable != "trust_scien") |> # remove science
  mutate(across(where(is.numeric), ~na_if(., 0))) # convert to NA
```


### Now plot all UK trends first

First create a data subset, this will be used later to compare vs other countries too 

```{r}
gb_trust_summaries <- trust_summaries_df |> filter(cntry == "GB")

gb_trust_summaries |> mutate(diff = ci_u - wt_avg_mean)
```


And plot - all trends for UK: 

```{r}
# Create the line chart with error bars
ggplot(gb_trust_summaries, 
       aes(x = essround, y = wt_avg_mean, 
           group = trust_variable, colour = trust_variable)) +  # group=1 ensures lines connect across points
  geom_line(size = 1) +  # Line for the trend
  geom_point(size = 1.5) +  # Points for each mean
  geom_errorbar(aes(ymin = ci_l, ymax = ci_u),  # Error bars using CI bounds
                width = 0.5,  # Width of the error bar caps
                color = "gray") +  
  labs(title = "Mean Trust in UK by variable, ESS round 1-11",
       subtitle = "With 95% Confidence Intervals",
       x = "ESS Round",
       y = "Mean Trust (/10 with CI included)") +
  theme_minimal()+
  
  # add extra labels to see each item
  geom_label_repel(
    data = . %>% group_by(trust_variable) %>% filter(essround == max(essround)),
    aes(label = trust_variable, fill = trust_variable), 
    color = "white",                # <--- CHANGE 1: Set text color inside labels to white
    fontface = "bold",
    alpha = 0.8,
    label.size = 0.5,
    show.legend = FALSE,
    box.padding = unit(0.5, "lines"),
    point.padding = unit(0.5, "lines"),
    segment.colour = NA,
    direction = "y",
    nudge_x = 2,
    min.segment.length = 0)+
    # set limits 
  scale_x_continuous(breaks = seq(1,11,1),
                     limits = c(min(gb_trust_summaries$essround), max(gb_trust_summaries$essround)+3)) +
  geom_vline(xintercept = 7.5, color = "orange", size = 1.5, alpha = 0.5)+
  theme(legend.position = "none")



```

Facet for UK only to have each vriable separate 
```{r}
# And print faceted by country too
ggplot(gb_trust_summaries, 
       aes(x = factor(essround), y = wt_avg_mean, 
           group = trust_variable, colour = trust_variable)) +  # group=1 ensures lines connect across points
  geom_line(size = 1) +  # Line for the trend
  geom_point(size = 1.5) +  # Points for each mean
  geom_errorbar(aes(ymin = ci_l, ymax = ci_u),  # Error bars using CI bounds
                width = 0.5,  # Width of the error bar caps
                color = "gray") +  
  facet_wrap(~trust_variable)+
  labs(title = "Mean Trust in UK by variable, ESS round 1-11",
       subtitle = "With 95% Confidence Intervals",
       x = "ESS Round",
       y = "Mean Trust (/10 with CI included)") +
  theme_minimal() + 
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for readability
    legend.position = "none"
  )+
  geom_vline(xintercept = 7.5) # sample of brexit timing

```

### Remove trust in science (missing)

As seen above, these are missing: 

```{r}
ess_data <- ess_data |> 
  select(-trust_scien)
```

We keep R1 of political parties as we can't remove just the one variable. 

## Investigating polarised values of trust

Create variables for very high trust and very low trust, we will group as: 
0,1,2,3: "Low trust"
4,5,6: "Moderate trust 
7,8,9,10: "High trust"

With a second variable we recode as: 
0,1: "Extreme low trust"
9,10: "Extreme high trust"

```{r}
# proof of concept for function to classify high/low
ess_data |> 
  mutate(trust_hilow_people = factor(case_when(trust_people<4 ~ "Low",
                                        between(trust_people, 4,6) ~ "Moderate",
                                        between(trust_people, 7,10) ~ "High",
                                        TRUE ~ NA),
                                     levels = c("Low", "Moderate", "High", NA))) |> 
  group_by(trust_hilow_people) |> count()

# check why NA values included  
ess_data |> 
  mutate(trust_hilow_people = case_when(trust_people<4 ~ "Low",
                                        between(trust_people, 4,6) ~ "Moderate",
                                        between(trust_people, 7,10) ~ "High",
                                        TRUE ~ NA)) |> 
  select(essround, cntry, trust_people, trust_hilow_people) |> 
  filter(is.na(trust_hilow_people))

```

We see that the NA's are true NA's so we can continue with our calculations: 

```{r}
#function to create high/low variables for each: 
low_high_trust <- function(trust_var) {
  factor(
    case_when(
      trust_var < 4 ~ "Low",
      between(trust_var, 4, 6) ~ "Moderate",
      between(trust_var, 7, 10) ~ "High", # use range to avoid any poorly coded 77/88/99 values being included
      TRUE ~ NA #account for true NAs from no response
  ),
  levels = c("High","Moderate","Low",  NA))
}

# and function for extreme trust (top/bottom 2 values)
extreme_trust <- function(trust_var) {
  factor(
    case_when(
      trust_var < 2 ~ "Extreme low",
      between(trust_var, 2, 8) ~ "Moderate",
      between(trust_var, 9, 10) ~ "Extreme high",
      TRUE ~ NA
  ),
  levels = c("Extreme high", "Moderate",  "Extreme low")
  )
}

# now apply over our trust variables to get the trend in trust
ess_data <- ess_data |> 
  mutate(across(matches("^trust") & !matches("hilow|extreme"), low_high_trust, .names = "{.col}_hilow"),
         across(matches("^trust") & !matches("hilow|extreme"), extreme_trust, .names = "{.col}_extreme")) 

```

Now calculate the unweighted rates within each country and wave 

```{r}
# subset the data
gb_polarise_trust <- ess_data |> 
  filter(cntry == "GB") |> 
  select(essround, cntry, matches("_extreme|_hilow"))

head(gb_polarise_trust)
```

Plot high and low trust views for UK in bars:

```{r}
# pivot longer for ggploting
gb_polarise_trust |> 
  select(essround, ends_with("hilow"), -matches("scien")) |> 
  pivot_longer(cols = ends_with("hilow"), 
               names_to = "trust_type", 
               values_to = "trust_category") |> 
  drop_na() |> # remove science and R1 parties
    # then plot stacked bars
  ggplot(aes(x = factor(essround), 
             fill = trust_category)) + 
  geom_bar(position = "fill") +  # stack 100%
  scale_y_continuous(labels = scales::percent_format()) +  
  scale_fill_manual(values = c("Low" = "red", "High" = "blue", "Moderate" = "NA")) + # custom colours to hide moderate
  facet_wrap(~trust_type, ncol=4) +  # Separate plots
  labs(title = "High and Low trends in trust vars (0-3 = Low; 7-10 = High)",
       x= "ESS Round", y = "respondents %", fill = "Trust Level") +
  theme_minimal()

```

And with the extreme trust levels (0-1 or 9-10):

```{r}
# pivot longer for ggploting
gb_polarise_trust |> 
  select(essround, ends_with("extreme"), -matches("scien")) |> 
  pivot_longer(cols = ends_with("extreme"), 
               names_to = "trust_type", 
               values_to = "trust_category") |> 
  drop_na() |> # remove science and R1 parties
    # then plot stacked bars
  ggplot(aes(x = factor(essround), 
             fill = trust_category)) + 
  geom_bar(position = "fill") +  # stack to 100%
  scale_y_continuous(labels = scales::percent_format()) +  
  scale_fill_manual(values = c("Extreme low" = "red", "Extreme high" = "blue", "Moderate" = "NA")) +  # custom colours to hide moderate
  facet_wrap(~trust_type, ncol=4) +  # Separate plots 
  labs(title = "Extreme trends in trust vars (0-1 = Low; 9-10 = High)",
       x= "ESS Round", y = "respondents %", fill = "Trust Level") +
  theme_minimal()
```

#### Comparing local and european parliament trust:

```{r}
# Create the line chart with error bars
gb_trust_summaries |> 
  filter(trust_variable %in% c("trust_europ", "trust_parliament")) |> 
ggplot(aes(x = essround, y = wt_avg_mean, 
           group = trust_variable, colour = trust_variable)) +  # group=1 ensures lines connect across points
  geom_line(size = 1) +  # Line for the trend
  geom_point(size = 1.5) +  # Points for each mean
  geom_errorbar(aes(ymin = ci_l, ymax = ci_u),  # Error bars using CI bounds
                width = 0.5,  # Width of the error bar caps
                color = "gray") +  
  labs(title = "Mean Trust in Parliament - UK vs European, ESS round 1-11",
       subtitle = "With 95% Confidence Intervals",
       x = "ESS Round",
       y = "Mean Trust (/10 with CI included)") +
  theme_minimal()+
  
  # add extra labels to see each item
  geom_label_repel(
    data = . %>% group_by(trust_variable) %>% filter(essround == max(essround)),
    aes(label = trust_variable, fill = trust_variable), 
    color = "white",                # <--- CHANGE 1: Set text color inside labels to white
    fontface = "bold",
    alpha = 0.8,
    label.size = 0.5,
    show.legend = FALSE,
    box.padding = unit(0.5, "lines"),
    point.padding = unit(0.5, "lines"),
    segment.colour = NA,
    direction = "y",
    nudge_x = 2,
    min.segment.length = 0)+
    # set limits 
  scale_x_continuous(breaks = seq(1,11,1),
                     limits = c(min(gb_trust_summaries$essround), max(gb_trust_summaries$essround)+3)) +
  geom_vline(xintercept = 7.5, color = "orange", size = 1.5, alpha = 0.5)+
  theme(legend.position = "none")


```


### International comparisons: 

```{r}
trust_summaries_df |> 
  ggplot(aes(x=essround,
             y=wt_avg_mean,
             group=cntry,
             colour = cntry))+
  geom_line()+
  # set limits 
  scale_x_continuous(breaks = seq(1,11,1),
                     limits = c(min(trust_summaries_df$essround), max(trust_summaries_df$essround)+1)) +
  theme_minimal()+
  geom_vline(xintercept = 7.5, color = "orange", size = 1.5, alpha = 0.5)+
  labs(title = "Average trust in [...] - ESS rounds 4 to 11, UK only",
     x = "ESS round", y = "Mean trust (0‒10)")+
  theme(legend.position = "bottom")+
  facet_wrap(~trust_variable)

```


#### Comparing individual countries

Matching in ESS_trust_w678.Rmd. The section is just for visual inspections. 

First, testing plots of just one country:

Create plot to test parallel trends between UK and other country. We write a function to complete the plotting against multiple countries more easily. 

```{r}
## print sample for comparing by country. 
trust_summaries_df |> 
  filter(cntry %in% c("GB", "DE")) |> 
  ggplot(aes(x = factor(essround), # Treat essround as a factor for discrete x-axis labels
             y = wt_avg_mean,
             group = cntry, # Group by country to draw lines for each country
             colour = cntry)) +
  geom_line(linewidth = 1) + # Added linewidth for better visibility
  geom_point(size = 2) + # Add points for each data point
  geom_vline(xintercept = 7.5, color = "orange", size = 1.5, alpha = 0.5)+
  facet_wrap(~trust_variable) +
  theme_minimal()+
  theme(legend.position = "bottom")+
  labs(title = "Average trust in [...] - ESS Round 1-11, GB vs DE",
   x = "ESS round", 
   y = "Mean trust (0‒10)",
   colour = "Country:")
```


Now make the function to plot GB vs other countries 
```{r}
# now create as function:
plot_gb_other <- function(summary_stats= summary_stats, other_country){
  summary_stats |> 
    select(essround, cntry, 
         people = ppltrst_mean, 
         euro_parl = trstep_mean, 
         legal_sys = trstlgl_mean,
         police = trstplc_mean,
         politicians = trstplt_mean,
         parliament = trstprl_mean,
         polit_parties = trstprt_mean,
         united_nats = trstun_mean
         ) |> 
  pivot_longer(cols = people:united_nats) |> 
  filter(cntry %in% c("GB", other_country),
         essround %in% c(6,7)) |> 
  ggplot(aes(x = factor(essround), # Treat essround as a factor for discrete x-axis labels
             y = value,
             group = cntry, # Group by country to draw lines for each country
             colour = cntry)) +
  geom_line(linewidth = 1) + # Added linewidth for better visibility
  geom_point(size = 2) + # Add points for each data point
  geom_vline(xintercept = 7.5, color = "orange", size = 1.5, alpha = 0.5)+
  facet_wrap(~name, scales = "free_y") +
  theme_minimal()+
  theme(legend.position = "bottom")+
  labs(title = glue("Average trust in [...] - ESS Round 6-7 (baseline), GB vs {other_country}"),
   x = "ESS round", 
   y = "Mean trust (0‒10)",
   colour = "Country:")
}

trust_summaries_df |> 
  filter(cntry %in% c("GB", "DE")) |> 
  ggplot(aes(x = factor(essround), # Treat essround as a factor for discrete x-axis labels
             y = wt_avg_mean,
             group = cntry, # Group by country to draw lines for each country
             colour = cntry)) +
  geom_line(linewidth = 1) + # Added linewidth for better visibility
  geom_point(size = 2) + # Add points for each data point
  geom_vline(xintercept = 7.5, color = "orange", size = 1.5, alpha = 0.5)+
  facet_wrap(~trust_variable) +
  theme_minimal()+
  theme(legend.position = "bottom")+
  labs(title = "Average trust in [...] - ESS Round 1-11, GB vs DE",
   x = "ESS round", 
   y = "Mean trust (0‒10)",
   colour = "Country:")


# now create as function:
plot_gb_other_parliament <- function(trust_summaries_df=trust_summaries_df, other_country){
  trust_summaries_df |> 
    filter(cntry %in% c("GB", other_country),
           trust_variable %in% c("trust_europ", "trust_parliament")) |> 
  ggplot(aes(x = factor(essround), # Treat essround as a factor for discrete x-axis labels
             y = wt_avg_mean,
             group = cntry, # Group by country to draw lines for each country
             colour = cntry)) +
  geom_line(linewidth = 1) + # Added linewidth for better visibility
  geom_point(size = 2) + # Add points for each data point
  geom_vline(xintercept = 7.5, color = "orange", size = 1.5, alpha = 0.5)+
  facet_wrap(~trust_variable, scales = "free_y") +
  theme_minimal()+
  theme(legend.position = "bottom")+
  labs(title = glue("Average trust in EU and local Parliament - ESS Round 1-11, GB vs {other_country}"),
   x = "ESS round", 
   y = "Mean trust (0‒10)",
   colour = "Country:")
}

plot_gb_other_parliament("DE")
```

Test plot:

```{r}
trust_summaries_df |> 
  filter(trust_variable %in% c("trust_europ", "trust_parliament"),
         cntry %in% c("GB", "DE")) |> 
ggplot(aes(x = essround, y = wt_avg_mean, 
           group = cntry, colour = cntry)) +  
  geom_line(size = 1) +  # Line for the trend
  geom_point(size = 1.5) +  # Points for each mean
  geom_errorbar(aes(ymin = ci_l, ymax = ci_u),  # Error bars using CI bounds
                width = 0.5,  # Width of the error bar caps
                color = "gray") +  
  labs(title = "Mean Trust in Parliament - UK vs European, ESS round 1-11",
       subtitle = "With 95% Confidence Intervals",
       x = "ESS Round",
       y = "Mean Trust (/10 with CI included)") +
  theme_minimal()+
  
  # add extra labels to see each item
  geom_label_repel(
    data = . %>% group_by(trust_variable) %>% filter(essround == max(essround)),
    aes(label = cntry, fill = cntry), 
    color = "white",    
    fontface = "bold",
    alpha = 0.8,
    label.size = 0.5,
    show.legend = FALSE,
    box.padding = unit(0.5, "lines"),
    point.padding = unit(0.5, "lines"),
    segment.colour = NA,
    direction = "y",
    nudge_x = 2,
    min.segment.length = 0)+
    # set limits 
  scale_x_continuous(breaks = seq(1,11,1),
                     limits = c(min(trust_summaries_df$essround), max(trust_summaries_df$essround)+3)) +
  geom_vline(xintercept = 7.5, color = "orange", size = 1.5, alpha = 0.5)+
  theme(legend.position = "none")+
  facet_wrap(~trust_variable)


```

Now turn to function to compare other countries 

```{r}
plot_gb_other_parliament <- function(trust_summaries_df=trust_summaries_df, other_country){
    
  trust_summaries_df |> 
      filter(trust_variable %in% c("trust_europ", "trust_parliament"),
             cntry %in% c("GB", other_country)) |> 
    ggplot(aes(x = essround, y = wt_avg_mean, 
               group = cntry, colour = cntry)) +  
      geom_line(size = 1) +  # Line for the trend
      geom_point(size = 1.5) +  # Points for each mean
      geom_errorbar(aes(ymin = ci_l, ymax = ci_u),  # Error bars using CI bounds
                    width = 0.5,  # Width of the error bar caps
                    color = "gray") +  
      labs(title =  glue("Average trust in EU and local Parliament - ESS Round 1-11, GB vs {other_country}"),
           subtitle = "With 95% Confidence Intervals",
           x = "ESS Round",
           y = "Mean Trust") +
      theme_minimal()+
  
  # add extra labels to see each item
  geom_label_repel(
    data = . %>% group_by(trust_variable) %>% filter(essround == max(essround)),
    aes(label = cntry, fill = cntry), 
    color = "white",    
    fontface = "bold",
    alpha = 0.8,
    label.size = 0.5,
    show.legend = FALSE,
    box.padding = unit(0.5, "lines"),
    point.padding = unit(0.5, "lines"),
    segment.colour = NA,
    direction = "y",
    nudge_x = 2,
    min.segment.length = 0)+
    # set limits 
  scale_x_continuous(breaks = seq(1,11,1),
                     limits = c(min(trust_summaries_df$essround), max(trust_summaries_df$essround)+3)) +
  geom_vline(xintercept = 7.5, color = "orange", size = 1.5, alpha = 0.5)+
  theme(legend.position = "none")+
  facet_wrap(~trust_variable)
}

```

## Compare parliament outcomes for GB and individual countries:
Here we can compare countries to look at comparability for parallel trends estimates. 

Countries in the dataset: 

```{r}
ess_data |> group_by(cntry, country_name) |> count()

```


Germany (DE):

```{r}
plot_gb_other_parliament(trust_summaries_df, "DE")
et```

Ireland (IE):

```{r}
plot_gb_other_parliament(trust_summaries_df, "IE")

```

Belgium (BE):
- doens't look feasible for parallel trends. 
```{r}
plot_gb_other_parliament(trust_summaries_df, "BE")
```

Switzerland:
```{r}
plot_gb_other_parliament(trust_summaries_df, "CH")
```

France:
- Also doesn't look promising. Trust in local parliament is massively differnt in the lead up to brexit. 
```{r}
plot_gb_other_parliament(trust_summaries_df, "FR")
```


Netherlands: 

```{r}
plot_gb_other_parliament(trust_summaries_df, "NL")

```

Scandi countries: 
The trends appear too positive pre-Brexit ot be comaprable. 
```{r}
plot_gb_other_parliament(trust_summaries_df, c("FI", "NO", "SE"))

```

Remaining EU countries: 
Not many promiising options here! 
```{r}
plot_gb_other_parliament(trust_summaries_df, c("CZ", "ES", "HU", "IL"))
plot_gb_other_parliament(trust_summaries_df, c("PO", "PT", "SI", "EE", "LT"))
```

Germany appears the best to satisfy parallel trends for a comparison of data. 

